---
title: "P8105_hw2_dxy2101"
author: "Daisy Yan"
date: "2022-10-04"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

### Load, clean, and use NYC transit dataset

```{r clean, message=FALSE}
subway_data = 
  read_csv(file = '~/Desktop/Data Science I/Homeworks/dxy2101_hw2_P8105/subway.csv') %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(
    entry = recode(entry, "YES" = TRUE, "NO" = FALSE)
    )
```

```{r name}
# List variables in dataset
names(subway_data)
```

The NYC transit dataset includes the following variables: line, station name, station latitude, station longitude, route numbers 1-11, entrance type, entry, vending, and ADA compliance.

Thus far, data cleaning has comprised of making all of the variable names the same format, retaining relevant variables, and converting the `entry` variable into a logical variable for future analysis. There are `r nrow(subway_data)` rows and `r ncol(subway_data)` columns.

This data is not tidy.

```{r proportion, echo=FALSE}
# Number of stations without vending that do not allow entrance
x = nrow(
  subway_data %>% 
  filter(vending=='NO', entry==FALSE) %>% 
  distinct(station_name, line)
  )

# Number of stations without vending that do allow entrance
y = nrow(
  subway_data %>% 
  filter(vending=='NO', entry==TRUE) %>% 
  distinct(station_name, line)
  )
```

There are `r nrow(distinct(subway_data, station_name, line))` distinct stations. There are `r nrow(subway_data %>% filter(ada==TRUE) %>% distinct(station_name, line))` stations that are ADA compliant. `r round(y/x, 2)` of the station entrances/exits without vending allow entrance.

```{r reformat}
# Convert all route variables to characters
subway_data =
  subway_data %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
  )
  
# Make route number and route name distinct variables
subway_data =
  pivot_longer(
    subway_data,
    route1:route11,
    names_to = "number",
    names_prefix = "route",
    values_to = "name"
  )
```

Now, this data is tidy.

Of the stations that serve the A train, `r nrow(subway_data %>% filter(name=='A', ada==TRUE) %>% distinct(station_name, line))` are ADA compliant.

# Problem 2

### Load, clean, and use Trash Wheel dataset

```{r mr, message=FALSE}
# Read and clean Mr. Trash Wheel dataset
mr_data = 
  read_excel('~/Desktop/Data Science I/Homeworks/dxy2101_hw2_P8105/trash.xlsx', sheet = 1) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>%
  na.omit(mr_data) %>%
  mutate(
    sports_balls = as.integer(round(sports_balls, 0))
  )
```

```{r prof}
# Read and clean Professor Trash Wheel dataset
prof_data =
  read_excel('~/Desktop/Data Science I/Homeworks/dxy2101_hw2_P8105/trash.xlsx', sheet = 2) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>%
  na.omit(prof_data) %>%
  mutate(
    sports_balls = as.integer(round(sports_balls, 0)),
    dumpster = as.character(dumpster)
  )
```

```{r join, message=FALSE}
# Add column indicating trash collector
mr_data = 
  data.frame(append(mr_data, c(collector="Mr"), after=1))

prof_data =
  data.frame(append(prof_data, c(collector="Professor"), after=1))

# Join data frames
full_df = 
  full_join(mr_data, prof_data)
```

The `full_df` dataset contains data on the number of dumpsters and types of trash collected by both Mr. Trash Wheel and Professor Trash Wheel from 2014 to beginning of 2021. In total, there are `r nrow(full_df)` observations. Key variables include: collector, data collected, weight (tons), types of trash collected and amount collected, and number of homes powered.

```{r weight, echo=FALSE}
prof =
  full_df %>%
  filter(collector=="Professor")
```

The total weight of trash collected by Professor Trash Wheel was `r sum(prof$weight_tons)` tons.

```{r sports, echo=FALSE}
trash = 
  full_df %>%
  filter(collector=="Mr", year == 2020)
```

The total number of sports balls collected by Mr. Trash Wheel in 2020 was `r sum(trash$sports_balls)`.

# Problem 3

### Load, clean, and use FiveThirtyEight datasets

```{r pols-month, message=FALSE}
# Clean data in pols-month.csv
pols_data = 
  read_csv(file = '~/Desktop/Data Science I/Homeworks/dxy2101_hw2_P8105/fivethirtyeight_datasets/pols-month.csv') %>%
  janitor::clean_names() %>%
  mutate(mon = as.character(mon)) %>%
  separate(col=mon, into=c("year", "month", "day"), sep='-') %>%
  mutate(month = recode(month, "01" = "jan", "02" = "feb",
                "03" = "mar", "04" = "apr", "05" = "may",
                "06" = "jun", "07" = "jul", "08" = "aug",
                "09" = "sep", "10" = "oct", "11" = "nov",
                "12" = "dec"))

# Make dataset tidy. Add variable for presidents.
pols_tidy_data = 
  pivot_longer(
    pols_data,
    cols = starts_with("prez"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "value") %>%
  mutate(
    value = as.numeric(value)
  ) %>%
  select(-value, -day)
```

```{r snp, message=FALSE}
# Clean data in snp.csv
snp_data = 
  read_csv(file = '~/Desktop/Data Science I/Homeworks/dxy2101_hw2_P8105/fivethirtyeight_datasets/snp.csv') %>%
  janitor::clean_names() %>%
  separate(col=date, into=c("month", "day", "year"), sep='/') %>%
  mutate(month = recode(month, "1" = "jan", "2" = "feb",
          "3" = "mar", "4" = "apr", "5" = "may", "6" = "jun", 
          "7" = "jul", "8" = "aug", "9" = "sep", "10" = "oct", 
          "11" = "nov","12" = "dec"),
         year = recode(year, "00" = "2000", "01" = "2001",
          "02" = "2002", "03" = "2003", "04" = "2004",
          "05" = "2005", "06" = "2006", "07" = "2007",
          "08" = "2008", "09" = "2009", "10" = "2010",
          "11" = "2011", "12" = "2012", "13" = "2013",
          "14" = "2014", "15" = "2015", "50" = "1950",
          "51" = "1951", "52" = "1952", "53" = "1953",
          "54" = "1954", "55" = "1955","56" = "1956",
          "57" = "1957", "58" = "1958", "59" = "1959",
          "60" = "1960", "61" = "1961", "62" = "1962",
          "63" = "1963", "64" = "1964", "65" = "1965",
          "66" = "1966", "67" = "1967", "68" = "1968",
          "69" = "1969", "70" = "1970", "71" = "1971",
          "72" = "1972", "73" = "1973", "74" = "1974",
          "75" = "1975", "76" = "1976", "77" = "1977",
          "78" = "1978", "79" = "1979", "80" = "1980",
          "81" = "1981", "82" = "1982", "83" = "1983",
          "84" = "1984", "85" = "1985", "86" = "1986",
          "87" = "1987", "88" = "1988", "89" = "1989",
          "90" = "1990", "91" = "1991", "92" = "1992",
          "93" = "1993", "94" = "1994", "95" = "1995",
          "96" = "1996", "97" = "1997", "98" = "1998",
          "99" = "1999")) %>%
  relocate(year, .before = month)
```

```{r unemployment, message=FALSE}
# Clean data in unemployment.csv
unem_data = 
  read_csv(file = '~/Desktop/Data Science I/Homeworks/dxy2101_hw2_P8105/fivethirtyeight_datasets/unemployment.csv') %>%
  janitor::clean_names()

# Make dataset tidy
unem_tidy_data =
  pivot_longer(
    unem_data,
    jan:dec,
    names_to = "month",
    values_to = "percent_unemployed"
    ) %>%
  mutate(
    year = as.character(year)
  )
```

```{r full, message=FALSE}
pols_snp =
  full_join(pols_tidy_data, snp_data)

pols_snp_unemployment = 
  full_join(pols_snp, unem_tidy_data)
```

`pols-month.csv`, `snp.csv`, and `unemployment.csv` were imported, cleaned and made tidy to create the `pols_snp_unemployment` dataset containing observations and variables from all three input datasets. `pols-month.csv`contains information on the political standing of national politicians at a given time. `snp.csv` contains information on the S&P market index at a given time. `unemployment.csv` contains information on the percentage of unemployment at a given time. All input datasets were cleaned to ensure that the given time points were organized by year and by month.

The resulting dataset, `pols_snp_unemployment`, contains `r nrow(pols_snp_unemployment)` observations and `r ncol(pols_snp_unemployment)` variables. The range of years is (`r range(pols_snp_unemployment$year)`). The key variables included are for a given year and month are the following: number of Republican governors, number of Democrat governors, number of Republican senators, number of Democrat senators, number of Republican representatives, number of Democrat representatives, whether or not the president at the time was a Republican or Democrat, the S&P index during the day and closing, and the percentage of people unemployed.
